PM Mungbean meta-analysis

### 2020-06-22 02_Preliminary_analysis.Rmd
More grammar and Markdown syntax edits

Before we exclude non-DMI fungicides lets have a look at the efficacy of the other fungicides in comparison to the DMIs.
For this comparison we will focus on the proportional yield saved, relative to the no spray control.
The primary focus of this research is the effect of fungicide on mitigating yield loss due to disease.
We are using proportional yield, relative to the no spray control, to reduce the effect of location and seasonal influences.

@@

Now we have calculated proportional yield gains we can inspect the difference between the fungicide active ingredients

```{r Fungicide_x_nSprays_tble2, message=FALSE}
PM_MB_means %>%
   group_by(fungicide_ai, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(
      caption = "Fungicide effect on proportion of yield gain.",
      col.names = c(
         "Fungicide active ingredient",
         "Total fungicide applications",
         "n Treatments",
         " 2.5% quartile",
         "Median proportional grain yield",
         "Mean proportional grain yield",
         "97.5 % quartile"
      ),
      align = "c"
   ) %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   column_spec(c(2, 3), width = "1cm") %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")
```

@@

There is a small yield effect among the strobilurin fungicide mixes and the DMI fungicides. 
While there is some differences in the variation between tebuconazole and propiconazole, the medians are very similar. 
I think it is safe to combine these fungicides, with the same mode of action in the meta-analysis.
The follow up spray seems to show an increase in yield protection.
Interestingly sulphur sprays show good protection when sprayed twice.

@@

#### Fungicide Doses  
From this point on we will investigate only the DMI fungicides, tebuconazole and propiconazole and disregard all other fungicide modes-of-action.

@@

The approximate period when in-crop rainfall impacts the most on the mean harvest is between 14 - 67 days after planting.

@@

### Season range and first incidence
Mungbean is a summer crop and powdery mildew requires cool conditions to establish.
Therefore the disease can be avoided by planting early in summer or spring, avoiding cooler temperatures at the end of some seasons.
Can we visualise when PM is most likely to establish over the calendar year?

@@

# For some trials where planting date and Harvest date is missing lets estimate 
#     the season length by using the mean of all season lengths
season_dates[is.na(season_dates$season_length),"cropping_season"] <- "Estimated"

@@

# Plot
season_dates %>%
   ggplot() +
   geom_pointrange(aes(
      x = trial_ref,
      y = Disease_day,
      ymin = Planting_day,
      ymax = Harvest_day,
      colour = cropping_season
   )) +
   scale_y_continuous(limits = c(
      0,
      max(
         season_dates$Planting_day +
            season_dates$season_length,
         na.rm = TRUE
      )
   ),
   labels = axis_labels_date,
   breaks = first_day_month) +
   geom_point(aes(x = trial_ref, y = Disease_day)) +
   coord_flip() +
   scale_color_manual(values=c("black","grey70"))+
   ggtitle("Season length and when powdery mildew was first spotted\nin each respective trial")

@@

First lets check we have all the AUDPC observations.

@@

>How well does 'area under the disease progress curve' (AUDPC) predict grain yield?

and _Does spray_management scenario mitigate yield loss through reducing AUDPC?_

@@

Lets tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship. 
Let's do the plot again for each trial where there are no missing values.  
Note: The plots are ordered by mean yield of the trial.

@@

This sequence of graphs show that AUDPC and disease incidence tends not impact on yields if the season is a low yielding, ie lower than 1 tonne per hectare.  
The higher the yield the greater greater the influence on the disease on the yield especially if the crop is high yielding with AUDPC also being large enough to reduce yield.

>>>>

Before we exclude non-DMI fungicides let's have a look at the efficacy of the other fungicides in comparison to the DMIs.
For this comparison we will focus on the proportional yield saved, relative to the no spray control.
The primary focus of this research is the effect of fungicide on mitigating yield loss due to disease.
We are using proportional yield, relative to the no spray control, to reduce the effect of location and seasonal influences.

@@

Now we have calculated proportional yield savings we can inspect the difference between the fungicide active ingredients

```{r Fungicide_x_nSprays_tble2, message=FALSE}
PM_MB_means %>%
   group_by(fungicide_ai, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(
      caption = "Fungicide effect on proportion of yield gain.",
      col.names = c(
         "Fungicide active ingredient",
         "Total fungicide applications",
         "n Treatments",
         "2.5% quartile",
         "Median proportional grain yield",
         "Mean proportional grain yield",
         "97.5 % quartile"
      ),
      align = "c"
   ) %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   column_spec(c(2, 3), width = "1cm") %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")
```

@@

There is a small yield effect among the strobilurin fungicide mixes and the DMI fungicides. 
While there is some differences in the variation between tebuconazole and propiconazole, the medians are very similar. 
I think it is safe to combine these fungicides, with the same mode of action in the meta-analysis.
The follow-up spray seems to show an increase in yield protection.
Interestingly sulphur sprays show good protection when sprayed twice.

@@

#### Fungicide Doses  
From this point on we will investigate only the DMI fungicides, tebuconazole and propiconazole, and disregard all other fungicide modes-of-action.

@@

The approximate period when in-crop rainfall affects the most on the mean harvest is between 14 - 67 days after planting.

@@

### Season range and first incidence

Mungbean is a summer crop and powdery mildew requires cool conditions to establish.
Therefore the disease can be avoided by planting early in summer or spring, avoiding cooler temperatures at the end of some seasons.
Can we visualise when PM is most likely to establish over the calendar year?

@@

# For some trials where planting date and Harvest date is missing let's estimate 
#     the season length by using the mean of all season lengths
season_dates[is.na(season_dates$season_length),"cropping_season"] <- "Estimated"

@@

# Plot
season_dates %>%
   ggplot() +
   geom_pointrange(aes(
      x = trial_ref,
      y = Disease_day,
      ymin = Planting_day,
      ymax = Harvest_day,
      colour = cropping_season
   )) +
   scale_y_continuous(limits = c(
      0,
      max(
         season_dates$Planting_day +
            season_dates$season_length,
         na.rm = TRUE
      )
   ),
   labels = axis_labels_date,
   breaks = first_day_month) +
   geom_point(aes(x = trial_ref, y = Disease_day)) +
   coord_flip() +
   scale_color_manual(values=c("black","grey70"))+
   ggtitle("Season length and when powdery mildew was first spotted\nin each respective trial")
```

@@

First let's check we have all the AUDPC observations.

@@

      _How well does 'area under the disease progress curve' (AUDPC) predict grain yield?_

   and

      _Does spray_management scenario mitigate yield loss through reducing AUDPC?_

@@

let's tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship. 
Let's do the plot again for each trial where there are no missing values.  
Note: The plots are ordered by mean yield of the trial.

@@

This sequence of graphs show that AUDPC and disease incidence tends not affect yields if the season is a low yielding season, _i.e._ lower than 1 tonne per hectare.
The higher the yield, the greater greater the influence on the disease on the yield especially if the crop is high yielding with AUDPC also being large enough to reduce yield.
